Formalizing Stack Safety as a Security Property

### 2023-02-04 submission5/core.tex
Clearer point in related work
CP intro -- rewritten contributions -- some notes and questions

Fortunately, other mechanisms besides these should be
amenable to strong 
formal guarantees.  In particular, 
Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} have presented an
array of tag-based micro-policies~\cite{pump_oakland2015} for stack
safety that give strong guarantees.
Their most realistic micro-policy, called Lazy Tagging and Clearing (\(LTC\)),
makes an interesting performance trade-off: it allows function
activations to improperly write into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
it afterward.

Because memory does get corrupted, \(LTC\) would not fulfill an
adaptation of Georges et al.'s property to the tagged setting.  But
there is a useful sense in which \(LTC\) does enforce stack safety.
While it does not necessarily protect individual addresses on the
stack, it does prevent one function activation from interfering with
another's observable behavior.

We propose a formal characterization of stack safety based on the intuition
of protecting each function activation from other activations.
We use the formal tools of language-based
security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
properties describing
the {\em integrity} and {\em confidentiality} of the caller’s local state
and the callee's behavior during the callee's execution, plus the control-flow protection
of {\em well-bracketed control flow} (\(\wbcf\)) articulated by
Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
While our properties are motivated by the desire to specify \(LTC\) precisely, 
they are stated somewhat abstractly in hopes that they may be applicable
to other enforcement mechanisms.

The structure of our properties enables them to be easily extended\bcp{??} to characterize systems with
features such as callee-save registers, argument-passing on the stack,
and tail-call elimination. We likewise extend \(LTC\) to protect these features, and
test the extended micro-policy against our property. We find that, apart from the extensions, \(LTC\)
is flawed in a way that undermines integrity and confidentiality, and we correct
this flaw.

@@

\begin{itemize}
\item A novel characterization of stack safety as a collection of properties:
  confidentiality and integrity for callee and caller, and well-bracketed
  control-flow. These can characterize the security requirements of systems
  that support only some portion of stack safety, and are
  parameterized over a notion of external observation, so that they can characterize
  lazy enforcement.
\item An instantiation of these definitions in a realistic
  setting with features such as argument passing
  on the stack, callee-saves registers, and tail-call elimination.
  Our model is modular enough that handling these features is straightforward.
\item The validation, via property-based randomize testing, of tag-based
  stack-safety policies. We find that \emph{Lazy Tagging and Clearing}
  falls short of stack safety, and propose a fix.
\end{itemize}

>>>>

Other mechanisms besides these should also be amenable to strong
formal guarantees.  In particular, Roessler and DeHon
\cite{DBLP:conf/sp/RoesslerD18} have presented an array of tag-based
micro-policies~\cite{pump_oakland2015} for stack safety that offer
strong guarantees.  Their most realistic micro-policy, called Lazy
Tagging and Clearing (\(LTC\)), makes an interesting performance
trade-off: it allows function activations to improperly write into one
another's stack frames, but ensures that the owner of the corrupted
memory cannot access it afterward.

Under this policy, one function activation {\em can} corrupt another's
memory---just not in ways that do not affect observable behavior.
Therefore, \(LTC\) would not fulfill Georges et al.'s property
(appropriately adapted to the tagged setting).  But \(LTC\) does
arguably enforce stack safety.  A looser, more observational
definition of stack safety is needed to fit this situation.

We propose here a formal characterization of stack safety, based on
the intuition of protecting function activations from each other.
\bcp{Aren't other formal definitions based on the same intuition?}
We use the tools of language-based
security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
properties describing
the {\em integrity} and {\em confidentiality} of the caller’s local state
and the callee's behavior during the callee's execution, plus the
{\em well-bracketed control flow} (\(\wbcf\)) property articulated by
Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
While our properties are motivated by the desire to specify \(LTC\) precisely, 
they are stated abstractly in hopes that they may also apply
to other enforcement mechanisms.

We find, via property-based random testing, that \(LTC\) is flawed in
a way that undermines both integrity and confidentiality; after
correcting this flaw, the repaired \(LTC\) satisfies all our
properties.  Moreover, we show that our properties can be extended to
describe calling conventions with features such as callee-save
registers, argument-passing on the stack, and tail-call
elimination. We modify \(LTC\) to protect these features and apply
random testing to validate this extended protection mechanism against
the extended properties.

@@

\begin{itemize}
\item We give a novel characterization of stack safety as a collection
of properties: confidentiality and integrity for callee and caller,
plus well-bracketed control-flow. Different subsets of these
properties characterize the security guarantees offered by mechanisms
that support only some aspects of stack safety\bcp{where do we show
  this?}.  The properties they are parameterized over a notion of
external observation, allowing them to characterize lazy enforcement
mechanisms.
\item We instantiate\bcp{extend??} these definitions in a realistic
setting with features such as\bcp{``such as''? are there others?}
argument passing on the stack, callee-saves registers, and tail-call
elimination.\bcp{How about: We extend these core definitions to
  describe a realistic setting with argument passing on the stack,
  callee-saves registers, and tail-call elimination.}  Our model is
modular enough that handling these features is straightforward.
\bcp{Throughout, the way we talk about the ``core model'' and its
  intantiation / extension / enrichment with realistic features feels
  quite weasily.}
\item We validate tag-based stack-safety enforcement via
property-based randomize testing. We find that \emph{Lazy Tagging and
  Clearing} falls short of ensuring stack safety, propose a fix, and
validate it.
\end{itemize}
